---
title: 'UHI: sensitivity for SVF'
author: "Marieke Dirksen"
date: "10 januari 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load libs, include=FALSE}
library(caret)
library(data.table)
library(dplyr)
library(ggmap)
library(ggplot2)
library(leaflet)
library(lubridate)
library(mapview)
library(plyr)
library(raster)
library(reshape2)
library(rgdal)
library(rgeos)
library(tidyverse)
library(uhimax)

wd<-"D:/uhimax/"

svf_utrecht<-list.files(paste0(wd,"SVF_new_subset"),pattern=".grd",full.names = TRUE)
svf_fig<-svf_utrecht[[1]]

grn<-raster("D:/groenkaart_rivm/rivm_20170415_g_groenkaart_10m.tif")

buildings<-readOGR(dsn="D:/OpenStreetMap/shp",layer="buildings_sens_ext")
buildings_crop<-readOGR(dsn="D:/OpenStreetMap/shp",layer="buildings_city_center")

uhi<-readRDS(paste0(wd,"Rdata/uhimax_obs_calc.rds"))
uhi<-uhi[which(uhi$stn %in% "UTRECHT23"),]

summary(uhi$meteo)
```

## SVF for the Wunderground without buildings
```{r}
svf_grn<-readRDS("D:/uhimax/Wunderground/wunderground.rds")

svf<-stack(paste0(wd,"SVF_res/svf_res_WGS84.grd")) # sky view factor values from 1 to 50m for the whole domain
# svf<-projectRaster(svf,crs=crs(svf_grn))
sp.buf<-gBuffer(svf,width=250,byid=TRUE) #gBuffer(svf_grn,width=250,byid=TRUE)

svf_wun<-crop(raster(svf_fig),extent(sp.buf))

svf_wun<-projectRaster(svf_wun,crs=crs(buildings))
sp.buf<-spTransform(sp.buf,CRS=crs(buildings))
sp.buf$svf<-NA

for(i in 1:length(sp.buf$Station.ID)){
sp.IUTRECHT299<-sp.buf[sp.buf$Station.ID==sp.buf$Station.ID[i],]

buildings_wun<-crop(buildings,extent(sp.IUTRECHT299))

svf_wun_msk<-mask(svf_wun,sp.IUTRECHT299)
svf_wun_no_buildings<-mask(svf_wun_msk,buildings_wun,inverse=TRUE)
sp.buf$svf[i]<-cellStats(svf_wun_no_buildings, stat='mean')
print(sp.buf$svf[i])
}

#transform back to RD
# sp.buf<-spTransform(sp.buf,CRS=crs("+init=epsg:28992"))#throws an error
# saveRDS(sp.buf,file="D:/uhimax/Wunderground/wunderground_buf.rds")

```

## Overview of data and observations
```{r}
ext_RD<-extent(c(136320, 136850, 454970, 456172))
ext_RD<-as(ext_RD,"SpatialPolygons")
crs(ext_RD)<-CRS("+init=epsg:28992")

center_box<-data.frame(maxlat = 52.098,minlat = 52.08,maxlong = 5.13,minlong = 5.109, id="uhi max")
center_box<-transform(center_box, laby=(maxlat +minlat )/2, labx=(maxlong+minlong )/2)

sens_box<-data.frame(maxlat = 52.1372 ,minlat = 52.05591  ,maxlong = 5.109605,minlong = 5.256057, id="svf sensitivity")
sens_box<-transform(sens_box, laby=(maxlat +minlat )/2, labx=(maxlong+minlong )/2)
sbbox <- make_bbox(lon = c(5.1,5.26), lat = c(52,52.14), f = .1)

```

## Gridded data from SVF and vegetation fraction 

Plotting all the data without excluding values. Extent of the area is cropped to the city center of Utrecht. 
```{r gridded data}
#does not seem to work with mapview
lopt = labelOptions(noHide = TRUE,
                    direction = 'top',
                    textOnly = TRUE,
                    style = list(
                                  "color" = "red",
                                  "font-family" = "serif",
                                  "font-style" = "italic",
                                  "box-shadow" = "3px 3px rgba(0,0,0,0.25)",
                                  "font-size" = "12px",
                                  "border-color" = "rgba(0,0,0,0.5)"
))

svf_grn<-readRDS(paste0(wd,"/Rdata/wunderground_svf_grn.rds"))

df.svf_grn<-data.frame(svf_grn)
df.svf_grn<-df.svf_grn[df.svf_grn$Station.ID %in% c("IUTRECHT376","IUTRECHT299"),]
df.svf_grn<-subset(df.svf_grn,select=c("Station.ID","Lon","Lat"))
sp.IUTRECHT<-df.svf_grn
coordinates(sp.IUTRECHT)<-~Lon+Lat
crs(sp.IUTRECHT)<-crs("+init=epsg:28992")

IUtrecht23<-data.frame(lon=52.079,lat=5.139)
coordinates(IUtrecht23)<- ~lat+lon
crs(IUtrecht23)<-crs("+init=epsg:4326")
IUtrecht23<-spTransform(IUtrecht23,crs(sp.IUTRECHT))

df.Wunderground<-data.frame(sp.IUTRECHT)
df.Wunderground$Station.ID<-as.character(df.Wunderground$Station.ID)
IUtrecht23<-c("IUtrecht23","137985.5","454554.1","TRUE")
df.Wunderground<-rbind(df.Wunderground,IUtrecht23)
df.Wunderground$Lon<-as.numeric(df.Wunderground$Lon)
df.Wunderground$Lat<-as.numeric(df.Wunderground$Lat)
coordinates(df.Wunderground)<-~Lon+Lat
crs(df.Wunderground)<-crs("+init=epsg:28992")
df.Wunderground<-spTransform(df.Wunderground,crs("+init=epsg:4326"))

#5.11461
# city_center<-extent(c(5.10961,5.13,52.08,52.098))
city_center<-extent(c(5.11461,5.13,52.08,52.098))

city_center<-as(city_center,"SpatialPolygons")
crs(city_center)<-CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")
city_center<-spTransform(city_center,CRS=crs("+init=epsg:28992"))
city_center_bf<-buffer(city_center,width=250)

sens_ext<-extent(c(5.109605,5.256057,52.05591,52.1372))
sens_ext<-as(sens_ext,"SpatialPolygons")
crs(sens_ext)<-CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")

map_ext<-extent(c(5.08,5.28,51.98,52.15))
map_ext<-as(map_ext,"SpatialPolygons")
crs(map_ext)<-CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")

# library(leaflet)
#export, width 450 height 450, crop top bottom to remove controlers
l.osm<-leaflet(df.Wunderground,options = leafletOptions(zoomControl = FALSE)) %>% addProviderTiles("OpenStreetMap") %>%
  addExtent(city_center,color="red",opacity=0.7,fillOpacity=0) %>%
  addExtent(sens_ext,color="blue",opacity=0.7,fillOpacity=0.1) %>% addScaleBar(position = "topleft") %>% addCircleMarkers(color=c("lightgreen","yellow","orange"),radius=4,opacity = 1,fillOpacity = 0.7) %>%  addSimpleGraticule(interval = 0.05) 
l.osm

l.esri<-leaflet(df.Wunderground,options = leafletOptions(zoomControl = FALSE)) %>% addProviderTiles("Esri.WorldImagery") %>%
  addExtent(city_center,color="red",opacity=0.7,fillOpacity=0) %>%
  addExtent(sens_ext,color="blue",opacity=0.7,fillOpacity=0.1) %>% addScaleBar(position = "topleft") %>% addCircleMarkers(color=c("lightgreen","yellow","orange"),radius=4,opacity = 1,fillOpacity = 0.7)  %>%  addSimpleGraticule(interval = 0.05) 
l.esri

#width=250,height=440
l.center<-
leaflet(options = leafletOptions(zoomControl =FALSE)) %>%    addProviderTiles("Esri.WorldImagery") %>%
addExtent(city_center,color="red",opacity=0.7,fillOpacity=0) %>%
addExtent(ext_RD,color="yellow",opacity=0.8,fillOpacity=0.1) %>% 
addScaleBar(position = "topleft") %>%          
addSimpleGraticule(interval = 0.01) 

l.center
```

```{r smooth grids city center}
# roads<-readOGR(dsn="D:/OpenStreetMap/shp",layer="gis_osm_roads_free_1")
# city_center_wgs<-spTransform(city_center,CRS=crs(roads))
# roads_crop<-crop(roads,city_center_wgs)
# roads_crop<-spTransform(roads_crop,crs(grn))
# roads_sp<-list("sp.lines",roads_crop,col="black",pch=1,cex=1)


grn.fveg<-crop(grn,extent(city_center_bf))
grn.fveg.blur<-blur_svf_fveg(grn.fveg)/100
grn.fveg.blur<-projectRaster(grn.fveg.blur,crs=crs(city_center))
grn.fveg.blur<-crop(grn.fveg.blur,city_center)


m.veg<-mapview(grn.fveg.blur,
        col.regions=rev(terrain.colors(n=41)),layer.name="vegetation fraction",at=seq(0.23,0.43,0.005),
        map.types =c("Esri.WorldImagery")) +
      mapview(city_center,alpha.regions=0, color="red",lwd=2) 
# mapshot(m.veg,file="C:/Users/marie/OneDrive/Afbeeldingen/sensitivity_fveg_blur.tiff")

raster_utrecht<-raster(svf_fig)
raster_utrecht<-crop(raster_utrecht,extent(city_center_bf))

raster_utrecht.blur<-blur_svf_fveg(raster_utrecht)
raster_utrecht.blur<-crop(raster_utrecht.blur,city_center)

m.svf<-mapview(raster_utrecht.blur,
        col.regions=rev(terrain.colors(n=26)),layer.name="sky view factor",at=seq(0.46,0.71,0.01),
        map.types =c("Esri.WorldImagery")) +
      mapview(city_center,alpha.regions=0, color="red",lwd=2) 
# mapshot(m.svf,file="C:/Users/marie/OneDrive/Afbeeldingen/sensitivity_svf_blur.tiff")
# raster_utrecht<-mapply(blur_svf_fveg,raster_utrecht)
# raster_utrecht_center<-mapply(crop,raster_utrecht,MoreArgs=list(y=extent(city_center)))
# grd.fveg_center<-crop(grd.fveg,extent(raster_utrecht_center[[1]]))

# mapview(raster_utrecht_center[[1]]) + grd.fveg_center + raster_utrecht_center[[2]] + raster_utrecht_center[[3]] #The minimum values of the SVF are located elsewhere for the 1m and 3-5m resolution runs!!!
```

## Extract SVF and VEGf from wunderground locations
```{r}
sens_ext<-spTransform(sens_ext,crs("+init=epsg:28992"))
grn_svf_sens<-crop(grn,sens_ext)
svf_svf_sens<-crop(raster(svf_fig),sens_ext)

# buildings<-readOGR(dsn="D:/OpenStreetMap/shp",layer="gis_osm_buildings_a_free_1")
# 
# sens_ext<-spTransform(sens_ext,CRS=crs(buildings))
# buildings_crop<-crop(buildings,sens_ext)
# 
# writeOGR(obj=buildings_crop, dsn="D:/OpenStreetMap/shp", layer="buildings_sens_ext", driver="ESRI Shapefile")
#crs(buildings)
# extent(buildings)
# city_center_bf_WGS<-spTransform(city_center_bf,crs(buildings))
# extent(city_center_bf_WGS)
# buildings_crop<-crop(buildings,extent(city_center_bf_WGS))
# writeOGR(obj=buildings_crop, dsn="D:/OpenStreetMap/shp", layer="buildings_city_center", driver="ESRI Shapefile")
# mapview(buildings_crop)


svf_svf_sens<-crop(svf_svf_sens,city_center_bf)
crs(svf_svf_sens)
svf_svf_sens<-projectRaster(svf_svf_sens,crs = crs(buildings_crop))
rasterOptions(maxmemory = 1e+09)
svf_msk<-mask(svf_svf_sens,buildings_crop,inverse=TRUE)
mapview(svf_msk)
# writeRaster(svf_msk,filename = "D:/uhimax/SVF_res/svf_utrecht_building_mask.grd")

grn_svf_sens_blur<-blur_svf_fveg(grn_svf_sens)/100
svf_svf_sens_blur<-blur_svf_fveg(svf_svf_sens)

#df.Wunderground@data$fveg<-extract(grn_svf_sens_blur,df.Wunderground)
#df.Wunderground@data$svf<-extract(svf_svf_sens_blur,df.Wunderground)
```

## Calculating the UHI according to Theeuwes

The SVF is masked using building shapefiles from OpenStreetMap. The averaged SVF with a radius of 250m excluding NA values is calculated.  

```{r calc uhimax}
uhi_center_utrecht<-function(skyview){
  skyview<-mask(skyview,buildings_crop,inverse=TRUE)
  skyview<-blur_svf_fveg(skyview)
  skyview<-raster::resample(skyview,grn.fveg.blur,method="bilinear")
  grd.fveg_center<-crop(grn.fveg.blur,extent(skyview))
  skyview<-crop(skyview,extent(grd.fveg_center))
  
  # values(skyview)[values(skyview)<0.2 | values(skyview)>0.9] = NA
  # values(grd.fveg_center)[values(grd.fveg_center)<0 | values(grd.fveg_center)>0.5] = NA

  uhi_center<-(2-grd.fveg_center-skyview)*median(uhi$meteo)
  return(list("svf"=skyview,"fveg"=grd.fveg_center,"meteo"=median(uhi$meteo),"uhi"=uhi_center))
  }
raster_utrecht_center<-lapply(svf_utrecht,stack)
buildings_crop<-spTransform(buildings_crop,crs(raster_utrecht_center[[1]]))

uhi_center<-lapply(raster_utrecht_center,uhi_center_utrecht)

saveRDS(uhi_center,file="D:/uhimax/SVF_res/uhi_center.rds")
mapview(uhi_center[[1]]$uhi) + uhi_center[[2]]$uhi + uhi_center[[3]]$uhi +city_center


```

```{r visualization}
library(lattice)
trellis.par.set(sp.theme())
roads<-readOGR(dsn="D:/OpenStreetMap/shp",layer="gis_osm_roads_free_1")
# buildings<-readOGR(dsn="D:/OpenStreetMap/shp",layer="gis_osm_buildings_a_free_1")

city_center_wgs<-spTransform(city_center,CRS=crs(roads))
# buildings_crop<-crop(buildings,city_center_wgs)

roads_crop<-crop(roads,city_center_wgs)


water<-readOGR(dsn="D:/OpenStreetMap/shp",layer="gis_osm_water_a_free_1")

uhi_utrecht<-stack(uhi_center[[1]]$uhi,uhi_center[[2]]$uhi,uhi_center[[3]]$uhi)

water_crop<-crop(water,city_center_wgs)
water_crop<-spTransform(water_crop,crs(uhi_utrecht))
water_sp<-list("sp.polygons",water_crop,fill="blue",first=FALSE)


names(uhi_utrecht)<-c("res=1m","res=3m","res=5m")
roads_crop<-spTransform(roads_crop,crs(uhi_utrecht))
roads_sp<-list("sp.lines",roads_crop,col="black",pch=1,cex=1)

scale = list("SpatialPolygonsRescale", layout.scale.bar(), 
	offset = c(136400,456580), scale = 300, which=1,
	fill=c("cyan","green"),first=FALSE)
s_text0 <- list("sp.text", c(136400, 456600 + 40), "0",col="green" ,cex = 1, which = 1,first=FALSE)
s_text1 <- list("sp.text", c(136400 + 300, 456600 + 40),col="green" , "300m", cex = 1, which = 1,first=FALSE)

scale2 = list("SpatialPolygonsRescale", layout.scale.bar(), 
	offset = c(136400,456580), scale = 300, which=1,
	fill=c("black","red"),first=FALSE)
s2_text0 <- list("sp.text", c(136400, 456600 + 40), "0",col="red",cex = 1, which = 1,first=FALSE)
s2_text1 <- list("sp.text", c(136400 + 300, 456600 + 40),"300m",col="red", cex = 1, which = 1,first=FALSE)


scale4 = list("SpatialPolygonsRescale", layout.scale.bar(), 
	offset = c(136400,456580), scale = 300, which=4,
	fill=c("cyan","green"),first=FALSE)
s4_text0 <- list("sp.text", c(136400, 456600 + 40), "0",col="cyan",cex = 1, which = 4,first=FALSE)
s4_text1 <- list("sp.text", c(136400 + 300, 456600 + 40),"300m",col="cyan", cex = 1, which = 4,first=FALSE)


#Urban heat island plot for Utrecht center
# uhi_utrecht<-projectRaster(uhi_utrecht,crs=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"))

png(filename = "C:/Users/marie/OneDrive/Afbeeldingen/uhi_utrecht.jpg",
    width=2800,
    height=1800,
    res=300)
spplot(uhi_utrecht,
       col.regions=rev(heat.colors(n=100)),
       alpha.regions=0.8,
       names.attr=c("(a)","(b)","(c)"),
       sp.layout=list(roads_sp,water_sp,scale,s_text0,s_text1),
       scales=list(draw=TRUE)
	)
dev.off()


#Green fraction in Utrecht Center
grn_svf_sens_blur<-resample(grn_svf_sens_blur,grn_svf_sens)
grn_svf_sens_utrecht<-crop(grn_svf_sens,city_center)/100
grn_svf_sens_blur_utrecht<-crop(grn_svf_sens_blur,city_center)
st.grn<-stack(grn_svf_sens_utrecht,grn_svf_sens_blur_utrecht)
png(filename = "C:/Users/marie/OneDrive/Afbeeldingen/fveg_utrecht.jpg",
    width=1800,
    height=1550,
    res=300)
spplot(st.grn,
       col.regions=rev(terrain.colors(n=100)),
       alpha.regions=0.8,
       names.attr=c("(a)","(b)"),
       sp.layout=list(roads_sp,water_sp,scale2,s2_text0,s2_text1),
       scales=list(draw=TRUE)
	)
dev.off()

# svf_utrecht<-stack(uhi_center[[1]]$svf,uhi_center[[2]]$svf,uhi_center[[3]]$svf)
# svf_utrecht<-crop(svf_utrecht,city_center)
# svf_utrecht<-resample(svf_utrecht,raster_utrecht_center[[1]])
# raster_utrecht_center<-mapply(crop,raster_utrecht_center,MoreArgs = list(y=extent(city_center)))
# svf_3m<-resample(raster_utrecht_center[[2]],raster_utrecht_center[[1]])
# svf_5m<-resample(raster_utrecht_center[[3]],raster_utrecht_center[[1]])
# svf_utrecht<-stack(raster_utrecht_center[[1]],svf_3m,svf_5m,svf_utrecht)
# saveRDS(svf_utrecht,file="D:/uhimax/SVF_res/svf_utrecht_resampled.rds")

#Sky view factor in Utrecht center
svf_utrecht<-readRDS("D:/uhimax/SVF_res/svf_utrecht_resampled.rds")
png(filename = "C:/Users/marie/OneDrive/Afbeeldingen/svf_utrecht.jpg",
    width=3300,
    height=3800,
    res=300)
spplot(svf_utrecht,
       layout=c(3,2),
       col.regions=terrain.colors(n=100),
       alpha.regions=0.8,
       names.attr=c("(a)","(b)","(c)","(d)","(e)","(f)"),
       sp.layout=list(roads_sp,water_sp,scale2,s2_text0,s2_text1),
       scales=list(draw=TRUE)
	)
dev.off()

svf_RD<-stack("D:/uhimax/SVF_res/st_sens_RD.grd")


scale3 = list("SpatialPolygonsRescale", layout.scale.bar(), 
	offset = c(136550,456072), scale = 150, which=1,
	fill=c("black","red"),first=FALSE)
s3_text0 <- list("sp.text", c(136550, 456072 + 40), "0",col="red",cex = 1.4, which = 1,first=FALSE)
s3_text1 <- list("sp.text", c(136550 + 150, 456072 + 40),"150m",col="red", cex = 1.4, which = 1,first=FALSE)

#Radius and direction sensitivity
svf_RD<-crop(svf_RD,ext_RD)
png(filename = "C:/Users/marie/OneDrive/Afbeeldingen/svf_RD.jpg",
    width=3800,
    height=2400,
    res=300)
spplot(svf_RD,
       layout=c(4,1),
       col.regions=terrain.colors(n=100),
       alpha.regions=0.8,
       names.attr=c("(1)","(2)","(3)","(4)"),
       sp.layout=list(roads_sp,water_sp,scale3,s3_text0,s3_text1),
       scales=list(draw=TRUE)
	)
dev.off()




```

```{r grid sens}
roads<-readOGR(dsn="D:/OpenStreetMap/shp",layer="gis_osm_roads_free_1")
grd_resolution<-stack("D:/uhimax/SVF_res/svf_res.grd")
sens_ext_wgs<-spTransform(sens_ext,CRS=crs(roads))
roads_crop<-crop(roads,sens_ext_wgs)
roads_crop<-spTransform(roads_crop,crs(grd_resolution))
roads_sp<-list("sp.lines",roads_crop,col="black",pch=1,cex=0.1,alpha=0.1)

water<-readOGR(dsn="D:/OpenStreetMap/shp",layer="gis_osm_water_a_free_1")
water_crop<-crop(water,sens_ext_wgs)
water_crop<-spTransform(water_crop,crs(grd_resolution))
water_sp<-list("sp.polygons",water_crop,fill="blue",col="blue",cex=0.2,alpha=0.3,first=FALSE)

scale5 = list("SpatialPolygonsRescale", layout.scale.bar(), 
	offset = c(137000,460000), scale = 2500, which=1,
	fill=c("black","red"),first=FALSE)
s5_text0 <- list("sp.text", c(137000, 460000 + 450), "0",col="red",cex = 1, which = 1,first=FALSE)
s5_text1 <- list("sp.text", c(137000 + 2500, 460000 + 450),"1.5km",col="red", cex = 1, which = 1,first=FALSE)

png(filename = "C:/Users/marie/OneDrive/Afbeeldingen/svf_res_sensitivity.jpg",
    width=2000,
    height=3300,
    res=300)
spplot(grd_resolution,
       layout=c(2,4),
       col.regions=terrain.colors(n=100),
       names.attr=c("(a)","(b)","(c)","(d)","(e)","(f)","(g)","(h)"),
       sp.layout=list(roads_sp,water_sp,scale5,s5_text0,s5_text1),
       scales=list(draw=TRUE))
dev.off()
```




## Correlations between the UHI calculations
```{r correlations}
corr1_3m<-corLocal(uhi_center[[1]]$uhi,uhi_center[[2]]$uhi)
corr3_5m<-corLocal(uhi_center[[2]]$uhi,uhi_center[[3]]$uhi)
corr1_5m<-corLocal(uhi_center[[1]]$uhi,uhi_center[[3]]$uhi)

mapview(corr1_3m) + # difference between the 1m and 3m
corr1_5m + # pearson corr 3m-5m
corr3_5m # 1-5m
```
