---
title: "Urban Heat Island Utrecht"
author: "Marieke Dirksen"
date: "22 november 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Initialization

The urban heat island equations are based on [Theeuwes (2017)](https://rmets.onlinelibrary.wiley.com/doi/full/10.1002/joc.4717).
```{r load libs, include=FALSE}
library(uhimax)
library(raster)
library(rgdal)
library(mapview)
library(data.table)
library(tidyverse)
library(lubridate)
library(dplyr)
library(caret)
```

## Initialization

```{r params}
wd<-"D:/uhimax/"

grd.svf<-stack(paste0(wd,"Grids_veg_svf/svf_utrecht_1m.grd"))
grd.fveg<-stack(paste0(wd,"Grids_veg_svf/greenness_utrecht_smooth_500m.grd"))
grd.fveg<-projectRaster(grd.fveg,crs=crs(grd.svf))

IUtrecht23<-data.frame(lon=52.079,lat=5.139)
coordinates(IUtrecht23)<- ~lat+lon
crs(IUtrecht23)<-crs("+init=epsg:4326")

mapview(grd.fveg) + mapview(IUtrecht23)

IUtrecht23_fveg<-raster::extract(grd.fveg,IUtrecht23)
IUtrecht23_svf<-raster::extract(grd.svf,IUtrecht23)
```

## From 10min data to Daily UHImax parameters

From the autmaomatic weather stations (AWS) in the rural area the following data is required:

* 10 minute global radiation
* 10 minute wind speed at 10m
* 10 minute temperature at 1.5m
* 10 minute pressure derived to MSL
* 1 hour relative humidity
* 1 hour precipitation 
* 1 hour mean wind

The 10 minute data from global radiation, wind speed, temperature and pressure is used to calculate the UHImax. The hourly relative humidity, precipitation and mean wind are used to filter out synoptic situations with frontal system and fog. Under these different conditions there is no clear relation between the AWS and city temperature.


```{r prepare Cabauw, warning=FALSE, cache=TRUE}
Cabauw_data<-prepare_Cabauw(ten_min = paste0(wd,"Cabauw_meteo/cabauw10min.csv"),
                            hourly_data = paste0(wd,"Cabauw_meteo/cabauw_hourly.csv"))
```

## Wunderground network data in Utrecht
Within the city of Utrecht we have carefully selected three stations from the [Wunderground network](www.wunderground.com):

* IUTRECHT23 (Used by [Theeuwes (2017)](https://rmets.onlinelibrary.wiley.com/doi/full/10.1002/joc.4717))
* IUTRECHT196
* IUTRECHT376
* IUTRECHT299

These stations have been measuring the city temperature for the past couple of years. The data can be downloaded using `download_time_seq`, note that the temperature is in Farenheid and not degrees Celcius. The data has been filtered and interpolated to 10minute timestamps. Unrealistic high and low temperatures are set to NA. In case more than 8 identical measurements in a row occur the value is also set to NA.

```{r other stations}
svf_grn<-readRDS(paste0(wd,"/Rdata/wunderground_svf_grn.rds"))
df.svf_grn<-data.frame(svf_grn)
df.svf_grn<-df.svf_grn[df.svf_grn$Station.ID %in% c("IUTRECHT196","IUTRECHT376","IUTRECHT299"),]
stations<-df.svf_grn$Station.ID
city_T<-paste0(wd,"Wunderground/Filtered/",df.svf_grn$Station.ID,"_filtered.rds")
svf<-df.svf_grn$svf
fveg<-df.svf_grn$grn

sp.IUTRECHT<-df.svf_grn
coordinates(sp.IUTRECHT)<-~Lon+Lat
crs(sp.IUTRECHT)<-crs("+init=epsg:28992")
mapview(grd.svf,layer.name="layer") + sp.IUTRECHT + IUtrecht23
```

## Calculate UHI max parameters 

Setting all parameters to required to calculate the UHI with the formula of [Theeuwes (2017)](https://rmets.onlinelibrary.wiley.com/doi/full/10.1002/joc.4717). For this figure no subset of meteorological conditions was made. 

```{r uhimax parameters,cache=TRUE}
uhi_Utrecht<-uhimax_params(STN="UTRECHT23",
              svf=IUtrecht23_svf,
              fveg=IUtrecht23_fveg,
              city_T=paste0(wd,"Wunderground/Filtered/IUTRECHT23_filtered.rds"),
              rural_T=Cabauw_data$Cabauw_10min,
              rural_meteo=Cabauw_data$Cabauw_Theeuwes)
ggplot(uhi_Utrecht,aes(Tref,Tcity,colour=S_new)) + geom_point() + geom_abline() + scale_color_gradientn(colours = heat.colors(20))

#Similarly the same function can be applied to other stations within Utrecht:
Wunderground_Utrecht<-mapply(uhimax_params, STN = stations, svf = svf, fveg = fveg, city_T = city_T,
                      MoreArgs = list(
                           rural_T = Cabauw_data$Cabauw_10min,
                           rural_meteo = Cabauw_data$Cabauw_Theeuwes
                           ,wd = paste0(wd,"UHImax/"), write.file = TRUE
                           ),SIMPLIFY = FALSE)
Wunderground_Utrecht<-do.call("rbind",Wunderground_Utrecht)

ggplot(Wunderground_Utrecht,aes(Tref,Tcity,colour=S_new)) + geom_point() + geom_abline() + scale_color_gradientn(colours = heat.colors(20))

```

## Subset based on weather conditions
```{r create a subset}
uhimax_files<-list.files(paste0(wd,"UHImax/"),
                         full.names = TRUE,pattern=".txt")
uhimax_files<-lapply(uhimax_files,fread)
uhimax_Utrecht<-do.call("rbind",uhimax_files)

# uhimax_Utrecht<-fread(paste0(wd,"UHImax/IUTRECHT23_UHIparams.txt"))

uhimax_Utrecht<-merge(Cabauw_data$Cabauw_sub,uhimax_Utrecht,by=c("start","stop"))
uhimax_Utrecht<-uhimax_Utrecht[which(uhimax_Utrecht$Rain==TRUE),]
uhimax_Utrecht<-uhimax_Utrecht[which(uhimax_Utrecht$Wind==TRUE),]
uhimax_Utrecht<-uhimax_Utrecht[which(uhimax_Utrecht$rh==TRUE),]
# uhimax_Utrecht<-uhimax_Utrecht[which(uhimax_Utrecht$Select==TRUE),]
uhimax_Utrecht<-uhimax_Utrecht[which(uhimax_Utrecht$Tref>17),]
# uhimax_Utrecht<-uhimax_Utrecht[which(uhimax_Utrecht$stn %in% c("IUTRECHT23","IUTRECHT376")),]
uhimax_Utrecht<-uhimax_Utrecht[which(uhimax_Utrecht$stn %in% c("IUTRECHT23","IUTRECHT196","IUTRECHT376","IUTRECHT299")),]
```

## Correlations with the three meteo params

The maximum urban heat island is in the cases we selected related to:

$$ UHI_{max} = (2-SVF-F_{veg}) \sqrt[4]{\frac{S*DTR^3}{U}} $$
The station `IUTRECHT23` has the best correlation coefficient with `0.6`, while also including the other observations in the center of Utrecht reduces R2 to `0.4`. 

```{r correlation UHImax}
uhi_u<-ggplot(uhimax_Utrecht,aes(U,UHImeasured,colour=DTR)) +geom_point() + scale_color_gradientn(colours = terrain.colors(20))
uhi_DTR<-ggplot(uhimax_Utrecht,aes(DTR,UHImeasured,colour=S_new)) + geom_point()+scale_color_gradientn(colours = heat.colors(20))
uhi_S<-ggplot(uhimax_Utrecht,aes(S_new,UHImeasured,colour=U)) +geom_point() + geom_point()+scale_color_gradientn(colours = topo.colors(20))
ggsave(uhi_u,filename = paste0(wd,"UHImax/fig/uhi_u.png"))
ggsave(uhi_DTR,filename = paste0(wd,"UHImax/fig/uhi_DTR.png"))
ggsave(uhi_S,filename = paste0(wd,"UHImax/fig/uhi_S.png"))
uhi_u
uhi_DTR
uhi_S

ggplot(uhimax_Utrecht,aes(UHImeasured,(2-svf-fveg)*meteo,colour=factor(stn))) +geom_point() +geom_abline() +xlim(0,10)+ylim(0,10)

caret::R2((2-uhimax_Utrecht$svf-uhimax_Utrecht$fveg)*uhimax_Utrecht$meteo,uhimax_Utrecht$UHImeasured)
caret::RMSE((2-uhimax_Utrecht$svf-uhimax_Utrecht$fveg)*uhimax_Utrecht$meteo,uhimax_Utrecht$UHImeasured)
```
## Using the formula of Theeuwes for Utrecht

The formula of Theeuwes(2017) was derived for:

* 0 < fveg < 0.4
* 0.2 < SVF < 0.9

Outside of this range the formula has not been tested.
```{r grid prections,eval=TRUE}
svf.r<-raster::resample(grd.svf,grd.fveg,method="bilinear")
values(svf.r)[values(svf.r)<0.2 | values(svf.r)>0.95] = NA
values(grd.fveg)[values(grd.fveg)<0 | values(grd.fveg)>0.4] = NA

city_center<-extent(c(5.1,5.13,52.08,52.098))
city_center<-raster(city_center)
values(city_center)<-1
crs(city_center)<-CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")
city_center<-projectRaster(city_center,crs=crs(grd.fveg))

st<-(2-svf.r-grd.fveg)*mean(uhimax_Utrecht$meteo)
st<-crop(st,city_center)
mapview(st)
```


<!-- ## Exploring non-linearity -->
<!-- ```{r visualization} -->
<!-- ggplot(uhimax_Utrecht,aes(UHImeasured,(2-svf-fveg)*meteo,colour=factor(stn))) +geom_point() +geom_abline() + xlim(0,10) +ylim(0,10) -->

<!-- ggplot(uhimax_Utrecht,aes(x=UHImeasured/(DTR*U),y=S/(rho*Cp*24))) +geom_point() +ylim(0,0.018) +xlim(0,0.80) -->

<!-- # y<-uhimax_Utrecht$S/(uhimax_Utrecht$rho*Cp*24) -->
<!-- # x<-uhimax_Utrecht$UHImeasured/(uhimax_Utrecht$DTR*uhimax_Utrecht$U) -->
<!-- # df<-data.frame(y,x) -->
<!-- # fit=nls(y ~ b*x^a,data=df,start=list(b=0.27,a=1.7)) -->

<!-- #p.meteo<-ggplot(uhimax_Utrecht,aes(UHImeasured,meteo,colour=factor(stn))) +geom_point() +geom_abline() -->
<!-- #p.meteo -->
<!-- #ggsave(p.meteo,filename = paste0(wd,"inst/UHImax/fig/uhimax_measured.png")) -->
<!-- ``` -->

<!-- ## Correlations with the vegetation and sky view factor -->
<!-- ```{r} -->
<!-- #####################Fitting it ourselfs -->
<!-- library(lme4) -->
<!-- fits<-lmList(formula=UHImeasured~0+meteo | stn,data=uhimax_Utrecht) #Residual standard error: 0.04697533 -->
<!-- coef_fits<-coef(fits) -->
<!-- names(coef_fits)<-"alpha" -->
<!-- coef_fits$stn<-rownames(coef_fits) -->
<!-- full_m<-merge(coef_fits,uhimax_Utrecht,by="stn") -->

<!-- # intercept_station<-coef_fits[1]+full_m$svf[1]+full_m$fveg[1] -->

<!-- # fit=lm(alpha ~ svf+fveg,data=full_m) -->
<!-- # a1=coef(fit)[1] -->
<!-- # a2=coef(fit)[2] -->
<!-- # a3=coef(fit)[3] -->

<!-- ggplot(data=full_m,aes(fveg,alpha))+geom_point()+ -->
<!--   xlab("fveg")+ -->
<!--   ylab("alpha")+ -->
<!--   geom_abline() -->

<!-- ggplot(data=full_m,aes(svf,alpha))+geom_point()+ -->
<!--   xlab("svf")+ -->
<!--   ylab("alpha")+ -->
<!--   geom_abline() -->

<!-- ggplot(data=full_m,aes(2-svf-fveg,alpha,colour=factor(stn)))+geom_point()+ -->
<!--   xlab("2-SVF-fveg")+ -->
<!--   ylab("alpha")+ -->
<!--   geom_abline() -->

<!-- ``` -->





